---
title: 'LM_IR_Proteomics'
author: Geremy Clair
date: March 1, 2021
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme : cerulean
    code_folding: hide
---

```{r knit_options, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=6, echo = TRUE, results = TRUE, warning = FALSE, message=FALSE) 
options(repos=structure(c(CRAN="https:/cloud.r-project.org/")))
setwd("~/Documents/GitHub/Effect_of_glomerular_disease_on_the_podocyte_cell_cycle/")
getwd()
```

# Required library

The required libraries are loaded 
- RomicsProcessor written by Geremy Clair (2022, in prep.) is used to perform traceable transformation and statistics to the dataset.
```{r load_packages}
library("RomicsProcessor")
```

# MaxQuant import

The LFQ data contained in the protein table was loaded, the corresponding metadata was loaded
```{r data_metadata_load}
data<-extractMaxQuant("~/Documents/05 - Lungmap/36 - LM_IR/01_source_files/proteinGroups.txt",quantification_type = "LFQ",cont.rm = T,site.rm = T,rev.rm = T)

data[,1]<-gsub("\\;.*","",data[,1])
data[,1]<-gsub(".*\\|","",data[,1])
representative_ID <- data[,1]

IDsdetails<-cbind(Identifiers=representative_ID, extractMaxQuantIDs("./01_source_files/proteinGroups.txt",cont.rm = T,site.rm = T,rev.rm = T))

metadata<- read.csv(file = "./01_source_files/metadata.csv")
write.csv(extractMaxQuantIDs("./01_source_files/proteinGroups.txt",cont.rm = T,site.rm = T,rev.rm = T),"./03_output_files/MaxQuantIDS.csv")
```

# Romics_object creation

The data and metadata were placed in an romics_object, the sample names were retrieved from the metadata, the condition  will be use for the coloring of the Figure and statistics
```{r create_romics}
romics_proteins<- romicsCreateObject(data,metadata,main_factor = "Condition")
```

# Full data analysis
## Data cleaning and normalization

The missingness was evaluated for each channel/sample
```{r missingness_eval_full}
romics_proteins<- romicsZeroToMissing(romics_proteins)
romicsPlotMissing(romics_proteins)
```
The data was log2 transformed, the distriution boxplot were then plotted
```{r log2_full}
romics_proteins<-log2transform(romics_proteins)
distribBoxplot(romics_proteins)
```

The proteins to be conserved for quantification were selected to contain at least 70% of complete values (3/4 samples) for a given condition, the overall missingness was evaluated after filtering.
```{r missing_filter_full}
romics_proteins<-romicsFilterMissing(romics_proteins,percentage_completeness = 60)
print(paste0(nrow(romics_proteins$data),"/", nrow(romics_proteins$original_data)," proteins remained after filtering", " (",round(nrow(romics_proteins$data)/nrow(romics_proteins$original_data)*100,2),"%)."))
romicsPlotMissing(romics_proteins)
```
As the same quantity of protein was labelled for each sample, the expectation is that the distribution of the protein abundance is centered, therefore a median centering was performed prior to plot again the distribution boxplots.
```{r median_norm_full}
romics_proteins<-medianCenterSample(romics_proteins)
distribBoxplot(romics_proteins)
```

## Data imputation

For some of the subsequent statistics imputations are required, we performed an imputation by assuming that the "non-detected" proteins were either low abundance or missing using the method developped by Tyranova et al. (PMID: 27348712). The gray distribution is the data distribution, the yellow distribution is the one for the random values used for imputation.
```{r imput_missing_full}
imputeMissingEval(romics_proteins,nb_stdev = 2,width_stdev = 0.5, bin=1)
romics_proteins<-imputeMissing(romics_proteins,nb_stdev = 2,width_stdev = 0.5)
```

## Statistics

The means and stdev are calculated for each group
```{r Means_stdevs_full}
romics_proteins<-romicsMean(romics_proteins)
romics_proteins<-romicsSd(romics_proteins)
```

ANOVA are performed
```{r statistics_ANOVA}
romics_proteins<-romicsANOVA(romics_proteins)
```

A heatmap depicting the proteins passing an ANOVA p<0.05 is plotted, the clusters obtained were saved in the statistics.
```{r Heatmap_full}
romicsHeatmap(romics_proteins,variable_hclust_number = 4,ANOVA_filter = "p", p=0.05,sample_hclust = F)
romics_proteins<-romicsVariableHclust(romics_proteins,clusters = 4,ANOVA_filter = "p",p= 0.05,plot = F)
romics_proteins<-romicsZscores(romics_proteins)
```

T.tests were added to the statistics
```{r ttest}
romics_proteins<-romicsTtest(romics_proteins,percentage_completeness = 60)
romicsVolcano(romics_proteins,p_type = "p",min_fold_change = 0)
```

The data was exported for further exploration and to generate supplementary dataset files.
```{r export_results}
results<-romicsExportData(romics_proteins,statistics = T,missing_data = T)
results<-cbind(Identifiers = rownames(results), results)
results<-merge(results,IDsdetails,by= "Identifiers")
write.csv(results, file= "./03_output_files/Data_processed_with_stats.csv")
save(romics_proteins,file = "03_output_files/romics_proteins.rda")
```